The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather

### 2023-05-03 Writing/manuscript_DR.Rmd
minor edits

---
title: The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather
author:
  - Taro Mieno^[Department of Agricultural Economics, University of Nebraska Lincoln, tmieno2@unl.edu]
  - Timothy Foster^[Department of WWW, University of Manchester, timothy.foster@manchester.ac.uk]
  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota]
  - Nicholas Brozovic^[Department of Agricultural Economics, University of Nebraska Lincoln, nbrozovic@nebraska.edu]
abstract: |
  Abstract goes here. It should be a single paragraph of around 200 words.
acknowledgments: |
  We wish to thank...
bibliography: DRA.bib
csl: iopart.csl
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
---

@@

Taking into account both impacts of saturated thickness, figure \@ref(fig:tot-impact) shows how water deficit affects yield at different saturated thickness levels evaluated. For corn, in years with water deficit of 400mm or less saturated thickness does not have economically significant negative impacts because the difference in irrigated and dryland yields are small and also because irrigated yields are similar irrespective of the level of saturated thickness. The same can be said for soybean except that the soybean is more tolerant than corn and the saturated thickness starts to make differences at water deficit of 500mm of higher. In the year with water deficit of 800mm, average corn yield at saturated thickness of 186 meters is about 15% and 25% higher than at 75.7 meters and 29.6 meters, respectively. These findings illustrate that vulnerability to drought is heterogeneous among irrigated production depending on the level of saturated thickness. In summary, our studies confirms that a decline in saturated thickness makes yield production much more vulnerable to severe drought, and its negative impact is non-negligible.

>>>>

---
title: The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather
author:
  - Taro Mieno^[Department of Agricultural Economics, University of Nebraska Lincoln, tmieno2@unl.edu]
  - Timothy Foster^[Department of WWW, University of Manchester, timothy.foster@manchester.ac.uk]
  - Shunkei Kakimoto^[Department of Applied Economics, University of Minnesota, kakim002@umn.edu]
  - Nicholas Brozovic^[Department of Agricultural Economics, University of Nebraska Lincoln, nbrozovic@nebraska.edu]
abstract: |
  Abstract goes here. It should be a single paragraph of around 200 words.
acknowledgments: |
  We wish to thank...
bibliography: DRA.bib
csl: iopart.csl
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
---

@@

```{r}
deficit_2012 <- data_corn[year == 2012, mean(balance)]

#--- corn ---#
balace_closest <- data_final_corn[which.min(abs(balance - deficit_2012)), balance]

corn_avg_yield_low <- data_final_corn[balance == balace_closest & sat_cat == "[39.4,82]", avg_yield]
corn_avg_yield_high <- data_final_corn[balance == balace_closest & sat_cat == "(194,708]", avg_yield]

corn_pct_decline <- (corn_avg_yield_high - corn_avg_yield_low) / corn_avg_yield_high * 100

#--- soy ---#
balace_closest <- data_final_soy[which.min(abs(balance - deficit_2012)), balance]

soy_avg_yield_low <- data_final_soy[balance == balace_closest & sat_cat == "[39.4,82]", avg_yield]
soy_avg_yield_high <- data_final_soy[balance == balace_closest & sat_cat == "(194,708]", avg_yield]

soy_pct_decline <- (soy_avg_yield_high - soy_avg_yield_low) / soy_avg_yield_high * 100
```

Taking into account both impacts of saturated thickness, figure \@ref(fig:tot-impact) shows how water deficit affects yield at different saturated thickness levels evaluated. For corn, in years with water deficit of 400mm or less saturated thickness does not have economically significant negative impacts because the difference in irrigated and dryland yields are small and also because irrigated yields are similar irrespective of the level of saturated thickness. The same can be said for soybean except that the soybean is more tolerant than corn and the saturated thickness starts to make differences at water deficit of 500mm of higher. In the year with water deficit of 950mm, average corn yield at saturated thickness of 186 meters is about 15% and 25% higher than at 75.7 meters and 29.6 meters, respectively. These findings illustrate that vulnerability to drought is heterogeneous among irrigated production depending on the level of saturated thickness. In summary, our studies confirms that a decline in saturated thickness makes yield production much more vulnerable to severe drought, and its negative impact is non-negligible.
